{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}

### 2012-05-10 defence/defence.tex
update defence for jiri; add livelock visual

\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic exploration versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

@@

	\begin{itemize}
		\item On exit, child's TCB is freed
		\item Forking thread does use-after-free
		\item Might return garbage instead of TID
	\end{itemize}
\end{frame}

@@

	{\bf Systematic Exploration} \related{Simsa '10}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	Previous work in kernel-space has not used systematic exploration.
	\linegap

@@

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
\end{frame}
\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{itemize}
		\item ``Everything interleaves with everything else''?
		\item What is a ``bug'', anyway?\related{Baumann '09}
	\end{itemize}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

\begin{frame}{Contribution} % slide: summary: "sys ex is a tool \ldots``
	Systematic exploration is a {\em tool} for testing concurrent systems.
	\begin{itemize}
		\item How can it be applied in kernel-space?
		\item What can we ask of the user to make it possible?
	\end{itemize}
	\linegap

@@

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	% TODO
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
	\linegap

	% TODO: does this slide want to be multicolumn, with an infinite loop diagram?
	Definite bugs
	\begin{itemize}
		\item Kernel panics / assertion failures
		\item Use-after-free
		\item Deadlock
	\end{itemize}
	Probable bugs
	\begin{itemize}
		\item Infinite loop
		\begin{itemize}
			\item Use structure of execution tree to judge stuckness.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

\begin{frame}{Studying Bugs In-Depth} % slide: visual explanation of bug
	% graphic want to be colored TODO
\end{frame}

@@

	Emphasis on ``steering'' of test parameters by experts
	\begin{itemize}
		\item Perhaps we can automate this steering?
	\end{itemize}
	\linegap
\end{frame}

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item Has potential to be a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}

>>>>

\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic testing versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

@@

	\begin{itemize}
		\item On exit, child's state is freed
		\item Forking thread does use-after-free
		\item Might return garbage instead of thread ID
	\end{itemize}
\end{frame}

@@

	{\bf Systematic Testing} \related{Musuvathi '08}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information, reproducibility
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	Previous work in kernel-space has not used systematic testing.
	\linegap

@@

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	% say: it's also important to understand the kernel's internals instead of just viewing it as one ``black box''
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
	% now that we see these internal interfaces, how do we apply systematic testing to it?
\end{frame}
\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{itemize}
		\item ``Everything interleaves with everything else''?
		\item What is a ``bug'', anyway?\related{Baumann '09}
	\end{itemize}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

\begin{frame}{Contribution} % slide: summary: "sys ex is a tool \ldots``
	Systematic testing is a {\em tool} for testing concurrent systems.
	\begin{itemize}
		\item How can it be applied in kernel-space?
		\item What simple things can we ask of the user to make it possible?
	\end{itemize}
	\linegap

@@

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	% Make sure to expound on what these are
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	% Say: ``combinatorial explosion''
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{Identifying Bugs}
	\begin{columns}[l]
		\column{0.75\textwidth}
		\textbf{How do we know we've found a bug?}
		\linegap

		Definite bugs
		\begin{itemize}
			\item Kernel panics / assertion failures
			\item Use-after-free
			\item Deadlock
		\end{itemize}
		Probable bugs
		\begin{itemize}
			\item Infinite loop
			\begin{itemize}
				\item Use structure of execution tree to judge progress.
			\end{itemize}
		\end{itemize}
		\column{0.25\textwidth}
		\includegraphics[width=\textwidth]{blank-livelock.png}
	\end{columns}
\end{frame}
\begin{frame}[fragile]{Identifying Bugs}
	\begin{columns}[l]
		\column{0.75\textwidth}
		\textbf{How do we know we've found a bug?}
		\linegap

		Definite bugs
		\begin{itemize}
			\item Kernel panics / assertion failures
			\item Use-after-free
			\item Deadlock
		\end{itemize}
		Probable bugs
		\begin{itemize}
			\item Infinite loop
			\begin{itemize}
				\item Use structure of execution tree to judge progress.
			\end{itemize}
		\end{itemize}
		\column{0.25\textwidth}
		\includegraphics[width=\textwidth]{livelock.png}
	\end{columns}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

@@

	Emphasis on users ``steering'' test by changing parameters
	\begin{itemize}
		\item Start with coarse granularity, iterate refining decision points
		\item Perhaps we can automate this steering?
	\end{itemize}
	\linegap
\end{frame}

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic testing in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item Has potential to be a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}
